Books and Babies: An Analysis of the Relationship between Female Literacy Rate and Fertility Rate in South Asia (2000-2020)

### 2024-04-19 paper/paper.qmd
Update paper.qmd

---
title: "Books and Babies: An Analysis of the Relationship between Female Literacy Rate and Fertility Rate in South Asia (2000-2020)"
author: "Jagpreet Singh"
thanks: "Code and data are available at: https://github.com/Jagpreet5ingh/female_fertility_rate_south_asia.git"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
bibliography: references.bib
execute: 
  echo: false
format: pdf
fig-pos: "H"
abstract: This paper utilizes data from the World Bank’s Gender Statistics database to examine the relationship between the adult female literacy rate (% of females ages 15 and above) and fertility rate (total births per woman) in South Asia between 2000-2020. The analysis revealed a negative correlation between the two variables, indicating that as the literacy rate increased, the fertility rate decreased. These findings matter as they highlight the importance of investing in female education and empowerment to manage population growth and promote sustainable development. The insights can guide policymakers make decisions on education initiatives and reproductive health programs to further enhance socioeconomic progress in the region.
---
```{r}
knitr::opts_chunk$set(echo = FALSE, results ="hide", message = FALSE, warning = FALSE)
```

# Introduction
According to the World Bank, the South Asian region comprises Afghanistan, Bangladesh, Bhutan, India, Maldives, Nepal, Pakistan, and Sri Lanka. It is the most populated region in the world with countries like India, Pakistan and Bangladesh which ranks among the top 10 in in terms of population size. Majority of these countries are also considered developing nations and play a major role in the world economy. In today’s globalized world where population dynamics significantly influence a country's course it is necessary to analyze gender statistics indicators such as fertility rate and female literacy rate and how they relate to each other. By doing so we can gain valuable insights which can help make policy decisions on female education and empowerment that inturn help manage population growth and promote sustainable development.

@@

# Model {#sec-model}

@@

$$ 
Yij = β0 + β1Xij + εij
$$
where:

@@

-   $\beta_1$ is the slope coefficient or the estimated change in Fertility Rate for a one-unit increase in the Female Literacy Rat.

@@

# Results {#sec-results}

@@

## Model Justification

# Discussion {#sec-discussion}

>>>>

---
title: "Books and Babies: An Analysis of the Relationship between Female Literacy Rate and Fertility Rate in South Asia (2000-2020)"
author: "Jagpreet Singh"
thanks: "Code and data are available at: https://github.com/Jagpreet5ingh/female_fertility_rate_south_asia.git"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
bibliography: references.bib
execute: 
  echo: false
format: pdf
number-sections: true
fig-pos: "H"
abstract: This paper utilizes data from the World Bank’s Gender Statistics database to examine the relationship between the adult female literacy rate (% of females ages 15 and above) and fertility rate (total births per woman) in South Asia between 2000-2020. The analysis revealed a negative correlation between the two variables, indicating that as the literacy rate increased, the fertility rate decreased. These findings matter as they highlight the importance of investing in female education and empowerment to manage population growth and promote sustainable development. The insights can guide policymakers make decisions on education initiatives and reproductive health programs to further enhance socioeconomic progress in the region.
---
```{r}
knitr::opts_chunk$set(echo = FALSE, results ="hide", message = FALSE, warning = FALSE)
```

# Introduction

According to the World Bank, the South Asian region comprises Afghanistan, Bangladesh, Bhutan, India, Maldives, Nepal, Pakistan, and Sri Lanka. It is the most populated region in the world with countries like India, Pakistan and Bangladesh which ranks among the top 10 in in terms of population size. Majority of these countries are also considered developing nations and play a major role in the world economy. In today’s globalized world where population dynamics significantly influence a country's course it is necessary to analyze gender statistics indicators such as fertility rate and female literacy rate and how they relate to each other. By doing so we can gain valuable insights which can help make policy decisions on female education and empowerment that inturn help manage population growth and promote sustainable development.

@@

# Model

@@

$$ 
Yij = β0 + β1Xij + εij
$$

where:

@@

-   $\beta_1$ is the slope coefficient or the estimated change in Fertility Rate for a one-unit increase in the Female Literacy Rate.

@@

# Results

@@

```
@tbl-model and Table 3 provides us with a vital summary of the regression model that further helps us to understand the relationship between fertility rate, total (births per woman) and literacy rate, adult females (% of females ages 15 and above). Together, these summary statistics tells us the following about our regression model:

* The t-value for the female literacy rate is approximately -19.811, which means that the estimate is -19.811 standard deviations away from 0.
The negative sign of the t-value indicates a negative relationship between female literacy rate and fertility rate in South Asia.

* The t-value for the intercept is 36.95598, indicating that the estimate is significantly different from zero.

* The p-value for both female literacy rate and the intercept is 0 which means that the coefficient for female literacy rate and the intercept is statistically significant at common alpha levels.

* The coefficient for the intercept is approximately 6.097, suggesting that when the female literacy rate is 0, the fertility rate is predicted to be around 6.097 births per woman. However, interpreting the intercept as meaningful may not be appropriate as a female literacy rate of 0 is not feasible and is outside the range of the data.

* The coefficient for female literacy rate is approximately -0.059, with a standard error of 0.003.This suggests that for each one-unit increase in the female literacy rate, the fertility rate is expected to decrease by 0.059 births per woman.
The negative coefficient indicates a negative relationship between female literacy rate and fertility rate which means as female literacy rate increases, the fertility rate will decrease.

* The R-squared value of 0.954 represents that approximately 95.4% of the variability in the fertility rate is explained by the model.This high R-squared value suggests a strong fit of the model to the data.

In summary, this linear model suggests a statistically significant negative relationship between female literacy rate and fertility rate which means as female literacy rate increases, the fertility rate will decrease. However, the intercept is not statistically significant. While the model has a high R-squared value, indicating a strong fit to the data, the practical significance of the intercept may be questionable.

# Discussion 
\newpage
# References
